---
title: "Multivariate ordinal response"
# author: Lu Mao (lmao@biostat.wisc.edu)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{poset_WR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates the win ratio (or net benefit) approach to multivariate ordinal data, under consensus or prioritized order between components.

## Data and Methodology

Let $Y = (y_1,\ldots, y_K)$ be a $K$-vector of ordinal responses, where without loss of generality we denote $y_k =0,1,\ldots, m_k -1$ 
These may be rating scores provided by different readers on the same subject's medical
image. The vector-valued response can be partially ordered by component-wise consensus, 
that is, we say $Y_i\prec Y_j$ if $Y_i\leq Y_j$ component-wise, with strict inequality
for at least one component. Alternatively, if some reader, say the first one, is more experienced than the rest, their score can be prioritized. Then we say 
$Y_i\prec Y_j$ if $y_{1i} < y_{1j}$, or $y_{1i} = y_{1j}$ and a consensus order
holds on the other components. In either case, we assume that a higher score is better,
so that $Y_i\succ Y_j$ (or $Y_j\prec Y_i$) means a "win" for $Y_i$ and a "loss" for $Y_j$.

### Two-group comparison
Let $Z = 1, 0$ be a binary indicator for treatment and control groups, respectively.
The probability of a treated winning against an untreated is
$P(Y_i\succ Y_j\mid Z_i = 1, Z_j = 0)$. Likewise, that of a treated winning against an untreated is

### Win ratio regression 

## Code and Example

### Basic syntax

### A data example


```{r setup}
library(poset)
```
